---
title: "R Notebook"
output: html_notebook
---

# Dataset: 
source: https://www.kaggle.com/piotrgrabo/breastcancerproteomes#clinical_data_breast_cancer.csv

```{r, libraries}
# load libraries
library(tidyverse)
```

```{r, setwd}
# change dir
setwd("/cloud/project/data/_raw/")
```

```{r, loadData}
# load the data
df_prot = read_csv(file = "77_cancer_proteomes_CPTAC_itraq.csv")
df_clin = read_csv("clinical_data_breast_cancer.csv")
df_PAM = read_csv("PAM50_proteins.csv")
```

Content:

File: 77cancerproteomesCPTACitraq.csv

* RefSeqaccessionnumber: RefSeq protein ID (each protein has a unique
ID in a RefSeq database)
* gene_symbol: a symbol unique to each gene (every protein is encoded
by some gene)
gene_name: a full name of that gene
* Remaining columns: log2 iTRAQ ratios for each sample (protein
expression data, most important), three last columns are from healthy
individuals


```{r}
# protein data: overview
glimpse(df_prot)

```
```{r}
patient_ID <- "([A-Z0-9]{2})[-](....)"
```

```{r}
df_prot %>%
  select_all(~str_replace(., pattern = patient_ID, "."))
```

